---
title: "Problem Set 3"
output: html_document
---
***
Diamonds:
How many observations?   = 53940
How many variables?      = 10
How many ordered factors? = 3
What value represent the best colour?   = D
***

```{r Load Data}
setwd("Documents/DataScience/SlideRule/EDA")
library(ggplot2)
data(diamonds)

str(diamonds)
```

***
Make histogram, describe the distributionincluding summary values 
such as the median and mean.

Data is skewed/long-tailed to the right, with values ranging from 326 to 18820. The median is 2401, which is lower than the mean of 3933, as we would expect in a long-tailed distribution. The middle 50% of prices fall between 950 and 5324.
***

```{r Diamonds: Histogram of Price}
ggplot(data = diamonds, aes(x = price)) +
      geom_histogram(color = 'black', fill = 'green') +
      xlab("Price") +
      ylab("Number of Diamonds")

summary(diamonds$price)
```

***
How many cost less than $500?
Less than $250?
$15000 or more?
***

```{r Diamonds: Counts by Cost}
length(subset(diamonds, diamonds$price < 500))  # doesn't work

table(diamonds$price<500)
table(diamonds$price<250)
table(diamonds$price>=15000)
```

***
Explore the largest peak in the
price histogram you created earlier.

Try limiting the x-axis, altering the bin width,
and setting different breaks on the x-axis.

```{r Diamonds: Exploring Histogram}
ggplot(data = diamonds, aes(x = price)) +
      geom_histogram(binwidth = 20) +
      scale_x_continuous(limits = c(0, 1000))

ggplot(data = diamonds, aes(x = price)) +
      geom_histogram(binwidth = 100, colour = 'black', fill = '#099DD9') +
      scale_x_continuous(breaks = seq(300, 1500, 100), limits = c(300, 1500))

```

***
Break out the histogram of diamond prices by cut. You should have five 
histograms in separate panels on your resulting plot.
***

```{r Diamonds: Price by Cut}
str(diamonds)

ggplot(data = diamonds, aes(x = price)) +
      geom_histogram(binwidth = 100, color = 'black', fill = 'green') +
      scale_x_continuous(breaks = seq(250, 10000, 100), limits = c(250, 10000)) +
      xlab("Price") +
      ylab("Number of Diamonds") +
      facet_wrap(~cut)

ggplot(data = diamonds, aes(x = price)) +
      geom_histogram(binwidth = 100, color = 'black', fill = 'green') +
      scale_x_continuous(breaks = seq(250, 5000, 100), limits = c(250, 5000)) +
      xlab("Price") +
      ylab("Number of Diamonds") +
      facet_wrap(~cut)

ggplot(data = diamonds, aes(x = price)) +
      geom_histogram(binwidth = 100, color = 'black', fill = 'green') +
      xlab("Price") +
      ylab("Number of Diamonds") +
      facet_wrap(~cut)

by(diamonds$price, diamonds$cut, summary)
by(diamonds$price, diamonds$cut, max)

```

***
Allow y-axis to vary between plots
***
```{r Diamonds: Free y-axis}
ggplot(data = diamonds, aes(x = price)) +
      geom_histogram(binwidth = 100, color = 'black', fill = 'green') +
      xlab("Price") +
      ylab("Number of Diamonds") +
      facet_wrap(~cut, scales="free_y")
```

***
price per carat
***

```{r Diamonds: Price per Carat}
ggplot(data = diamonds, aes(x = price/carat)) +
      geom_histogram(color = 'black', fill = "#009E73") +
      xlab("Price per Carat") +
      ylab("Number of Diamonds") +
      facet_wrap(~cut, scales="free_y") +
      scale_x_log10()

ggsave('pricepercarat.png')
```

***
Investigate the price of diamonds using box plots, numerical summaries, 
and one of the following categorical variables: cut, clarity, or color.
***

```{r Diamonds: Price by Cut}
str(diamonds)

ggplot(data = diamonds, aes(x = cut, y = price)) +
      geom_boxplot()

ggplot(data = diamonds, aes(x = cut, y = price)) +
      geom_boxplot() +
      coord_cartesian(ylim = c(250, 8000))

table(diamonds$cut)
by(diamonds$price, diamonds$cut, summary)

ggplot(data = diamonds, aes(x = color, y = price)) +
      geom_boxplot()

ggplot(data = diamonds, aes(x = cut, y = price)) +
      geom_boxplot() +
      coord_cartesian(ylim = c(250, 10000))

table(diamonds$color)
by(diamonds$price, diamonds$color, summary)

IQR(subset(diamonds, color == 'D')$price)
IQR(subset(diamonds, color == 'J')$price)
```

***
investigate price per carat across color, using boxplots
***

```{r Diamonds: Price per Carat by Colour}
ggplot(data = diamonds, aes(x = color, y = price/carat)) +
      geom_boxplot()

ggplot(data = diamonds, aes(x = color, y = price/carat)) +
      geom_boxplot() +
      coord_cartesian(ylim = c(0, 7500))

ggplot(data = diamonds, aes(y = price/carat, x = color)) +
      geom_boxplot() +
      facet_wrap(~color, scales = 'free_y')
```

***
Investigate the weight of diamonds using a freq polygon. Change binwidth to 
see how polygons are affected. What carat size has a count greter than 2000?
***

```{r Diamonds: Frequency Polygons}
ggplot(data = diamonds, aes(x = carat)) +
      geom_freqpoly()

ggplot(data = diamonds, aes(x = carat)) +
      geom_freqpoly(binwidth = .1) +
      scale_x_continuous(breaks = seq(0, 3, .1), limits = c(0, 3))

ggplot(data = diamonds, aes(x = carat)) +
      geom_freqpoly(binwidth = .01) +
      scale_x_continuous(breaks = seq(0, 3, .1), limits = c(0, 3))
```

***
GapMinder data EDA
data set = killed in storms
***

```{r GapMinder Data}
library(tidyr)
library(dplyr)
library(gapminder)

gapm <- gapminder
head(gapm)

by(gapm$lifeExp, gapm$continent, median)
aggregate(lifeExp ~ continent, gapm, median)

gapm %>%
    filter(year == 2007) %>%
    group_by(continent) %>%
    summarise(lifeExp = median(lifeExp))

ggplot(gapm, aes(x = pop)) +
      geom_histogram() +
      scale_x_log10()

ggplot(gapm, aes(x = lifeExp)) +
      geom_histogram() +
      theme_bw()

ggplot(gapm, aes(x = gdpPercap)) +
      geom_histogram() +
      theme_bw()

ggplot(gapm, aes(x = gdpPercap)) +
      geom_histogram() +
      scale_x_log10() +
      theme_bw()

ggplot(gapm, aes(x = continent, y = lifeExp)) +
      geom_boxplot(outlier.colour = "hotpink")

ggplot(gapm, aes(x = continent, y = lifeExp)) +
      geom_boxplot(outlier.colour = "hotpink") +
      geom_jitter(position = position_jitter(width = 0.1, height = 0), alpha = 1/4)

ggplot(gapm, aes(x = continent, y = pop)) +
      geom_boxplot(outlier.colour = "hotpink") +
      coord_cartesian(ylim = c(0, 250000000))

ggplot(gapm, aes(x = continent, y = gdpPercap)) +
      geom_boxplot(outlier.colour = "hotpink") +
      coord_cartesian(ylim = c(0, 40000))

highgdp <- subset(gapm, gdpPercap >= 30000)

ggplot(highgdp, aes(x = continent, y = gdpPercap)) +
      geom_boxplot()

```
