---
title: "problem_set_4"
output: html_document
---


```{r Scatterplot of price vs x}
setwd("Documents/DataScience/SlideRule/EDA")
library(ggplot2)

data(diamonds)
names(diamonds)

ggplot(diamonds, aes(x = x, y = price)) +
      geom_point()
```


```{r Correlations with price}

with(diamonds, cor.test(price, x))
with(diamonds, cor.test(price, y))
with(diamonds, cor.test(price, z))
```
price, x  r^2 = 0.88
price, y  r^2 = 0.87
price, z  r^2 = 0.86

***

Scatterplot of price v depth. Change the code to make the transparency of the
points to be 1/100 of what they are now and mark the x-axis every 2 units. Calculate correlation.

```{r Price v Depth}
ggplot(diamonds, aes(x = depth, y = price)) +
      geom_point()

max(diamonds$depth)
max(diamonds$price)

ggplot(diamonds, aes(x = depth, y = price)) +
      geom_point(alpha = 1/100) +
      scale_x_continuous(breaks = seq(0, 79, 2))

with(diamonds, cor.test(price, depth))
```

Most diamonds have a depth value between 58 and 64. Correlation of price with 
depth is -0.011.

***
Scatterplot of price v carat, ommitting top 1% of values.

```{r Price v Carat}
ggplot(diamonds, aes(x = carat, y = price)) +
      geom_point() +
      xlim(0, quantile(diamonds$carat, .99)) +
      ylim(0, quantile(diamonds$price, .99))

```

***
Create a scatterplot of price vs. volume (x * y * z). This is a very 
rough approximation for a diamond's volume.  Create a new variable 
for volume in the diamonds data frame. Run correlation on subset of 
volume.

```{r Price v volume}

ggplot(diamonds, aes(x = (x*y*z), y = price)) +
      geom_point() +
      xlab("volume")

diamonds$volume <- with(diamonds, x*y*z)

library(plyr)
count(diamonds$volume == 0)

detach("package:plyr", unload=TRUE)

with(subset(diamonds, volume >0 & volume <=800), cor.test(price, volume))

```

Observations: two possible outliers, one with a volume value near 800, 
another with a volume near 3800.  20 diamonds have volume = 0. otherwise 
distribution looks exponential. Correlation of price with volume, when 
volume is >0 and <=800, is 0.924.

***

Adjusted plot for price v volume. Add linear model. Is this a good model 
for these data?

```{r Price v Volume, adjusted}

ggplot(subset(diamonds, volume >0 & volume <=800), 
       aes(x = volume, y = price)) +
      geom_point(alpha = 1/10) +
      ylim(0, 20000) +
      geom_smooth(method = 'lm', colour = 'red')

```

Observations: Not the worst model, but not optimal.  The relationship between
price and volume appears to be exponential, not linear.

***

Create a new data frame containing info on diamonds by clarity. 
Name the data frame diamondsByClarity. It should contain the following 
variables in this order:
     (1) mean_price
     (2) median_price
     (3) min_price
     (4) max_price
     (5) n
where n is the number of diamonds in each level of clarity.

```{r Diamonds by Clarity}

library(dplyr)

diamondsByClarity <- diamonds %>%
      group_by(clarity) %>%
      summarize(mean_price = mean(price),
                median_price = median(price) * 1.0,   # to avoid error below
                min_price = min(price),
                max_price = max(price),
                n = n()) %>%
      arrange(clarity)

# error: loss of precision when attempting to convert numeric to integer
```

***

Bar charts, colour and clarity. Create two bar plots on one output 
image using the grid.arrange() function from the package gridExtra.

```{r Colour and Clarity}
library(gridExtra)

diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

p1 <- ggplot(diamonds_mp_by_clarity, aes(x = clarity, y = mean_price)) +
          geom_bar(stat = 'identity') +
          ylab("average price")
           

p2 <- ggplot(diamonds_mp_by_color, aes(x = color, y = mean_price)) +
      geom_bar(stat = 'identity') +
       ylab("average price")

grid.arrange(p1, p2)

diamonds_by_cut <- group_by(diamonds, cut)
diamonds_mp_by_cut <- summarise(diamonds_by_cut, mean_price = mean(price))

ggplot(diamonds_mp_by_cut, aes(x = cut, y = mean_price)) +
      geom_bar(stat = 'identity') +
       ylab("average price")

```

***
Gapminder con't: examine pairs of variable and create 2-5 plots that make
use of the techniques from Lesson 4.

```{r Gapminder EDA}
library(gapminder)

gapm <- gapminder
head(gapm)
names(gapm)
summary(gapm$year)

with(gapm, cor.test(lifeExp, gdpPercap))

gapm_1952_by_cont <- gapm %>%
      filter(year == 1952) %>%
      group_by(continent) %>%
      summarize(mean_lifeExp = mean(lifeExp),
                median_lifeExp = median(lifeExp),
                mean_pop = mean(pop),
                median_pop = median(pop),
                mean_gdp = mean(gdpPercap),
                median_gdp = median(gdpPercap),
                n = n())

gapm_2007_by_cont <- gapm %>%
      filter(year == 2007) %>%
      group_by(continent) %>%
      summarize(mean_lifeExp = mean(lifeExp),
                median_lifeExp = median(lifeExp),
                mean_pop = mean(pop),
                median_pop = median(pop),
                mean_gdp = mean(gdpPercap),
                median_gdp = median(gdpPercap),
                n = n())

ggplot(gapm_1952_by_cont, aes(x = continent, y = mean_lifeExp)) +
      geom_bar(stat = 'identity')

p1 <- ggplot(gapm_1952_by_cont,
             aes(x = continent, y = median_lifeExp, fill = continent)) +
      geom_bar(stat = 'identity') +
      scale_y_continuous(limits = c(0, 80)) +
      ylab("median life expectancy") +
      ggtitle("1952")

p2 <- ggplot(gapm_2007_by_cont, 
             aes(x = continent, y = median_lifeExp, fill = continent)) +
      geom_bar(stat = 'identity') +
      ylab("median life expectancy") +
      ggtitle("2007")

grid.arrange(p1, p2)

p3 <- ggplot(gapm_1952_by_cont,
             aes(x = continent, y = median_gdp, fill = continent)) +
      geom_bar(stat = 'identity') +
      scale_y_continuous(limits = c(0, 30000)) +
      ylab("median GDP") +
      ggtitle("1952")

p4 <- ggplot(gapm_2007_by_cont, 
             aes(x = continent, y = median_gdp, fill = continent)) +
      geom_bar(stat = 'identity') +
      ylab("median gdp") +
      ggtitle("2007")

grid.arrange(p3, p4)

ggplot(gapm, aes(x = gdpPercap, y = lifeExp, color = year)) +
      geom_point() +
      ylab("life expectancy") +
      xlab("GDP per capita")
```

