---
title: "Problem Set 5"
output: html_document
---


```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
setwd("Documents/DataScience/SlideRule/EDA")

data(diamonds)
```

***
Create a histogram of diamond prices.Facet the histogram by diamond color 
and use cut to color the histogram bars.

```{r Price Histograms with Facet and Color}
library(RColorBrewer)

ggplot(diamonds, aes(x = price, fill = cut)) +
      geom_histogram() +
      scale_x_log10() +
      facet_wrap(~ color) +
      scale_fill_brewer(type = 'qual')

```

***
Create a scatterplot of diamond price vs. table and color the points by 
the cut of the diamond.

```{r Price vs. Table Colored by Cut}

ggplot(diamonds, aes(x = table, y = price, color = cut)) +
      geom_jitter(size = 3, h = 0) +
      scale_x_continuous(breaks = seq(50, 80, 2), limits = c(50, 80)) +
      scale_color_brewer(type = 'qual')  # qualitative color scheme

```

***
Create a scatterplot of diamond price vs. volume (x * y * z) and color 
the points by the clarity of diamonds. Use scale on the y-axis to take the 
log10 of price. You should also omit the top 1% of diamond volumes from 
the plot.

```{r Price vs. Volume and Diamond Clarity}
diamonds$volume <- with(diamonds, x*y*z)
ggplot(subset(diamonds, volume > 0), 
       aes(x = volume, y = price, color = clarity)) +
      geom_point() +
      scale_y_log10() +
      xlim(0, quantile(diamonds$volume, .99)) +   # omit top 1% of values
      scale_color_brewer(type = 'div')   #diverging color scheme

```

***
Create a new variable called 'prop_initiated' in the Pseudo-Facebook 
data set. The variable should contain the proportion of friendships 
that the user initiated. Then create a line graph of the median proportion of
friendships initiated ('prop_initiated') vs. tenure and color the line 
segment by year_joined.bucket.

```{r Proportion of Friendships Initiated}
pf <- read.delim('pseudo_facebook.tsv')
names(pf)

pf$prop_initiated <- with(pf, round(friendships_initiated / friend_count * 100))
pf$year_joined <- floor(2014 - pf$tenure / 365)
pf$year_joined.bucket <- cut(pf$year_joined, 
                             breaks = c(2004, 2009, 2011, 2012, 2014),
                             include.lowest = FALSE)

# line graph of the median proportion of friendships initiated
# ('prop_initiated') vs. tenure, line colored by year_joined.bucket
ggplot(subset(pf, !is.na(prop_initiated)), aes(x = tenure, y = prop_initiated)) +
      geom_line(aes(colour = year_joined.bucket), stat = 'summary', fun.y = median)

# code above doesn't work, although it did work once. code below is same 
# structure and works. stackoverload says issue w R 3.2.0
ggplot(subset(pf, !is.na(gender)), aes(x = age, y = friend_count)) +
      geom_line(aes(colour = gender), stat = 'summary', fun.y = median)

# recreate graph with smoother
ggplot(subset(pf, !is.na(prop_initiated)), aes(x = tenure, y = prop_initiated)) +
      geom_smooth(aes(colour = year_joined.bucket))

# which group has highest proportion of friendships initiated? = 2012-2014
# what is their mean proportion initiated?   = 0.6654
by(pf$prop_initiated, pf$year_joined.bucket, summary)

```

***
Create a scatter plot of the price/carat ratio of diamonds. The variable 
x should be assigned to cut. The points should be colored by diamond color, 
and the plot should be faceted by clarity.

```{r Price/Carat Binned, Faceted, & Colored}
data(diamonds)
ggplot(diamonds, aes(x = cut, y = price / carat, color = color)) +
      geom_jitter() +
      facet_wrap(~ clarity) +
      scale_color_brewer(type = 'div')

```

***
Gapminder con't
In your investigation, examine 3 or more variables and create 2-5 plots 
that make use of the techniques from Lesson 5.

```{r Gapminder}
library(gapminder)

gapm <- gapminder
head(gapm)
names(gapm)
table(gapm$year)

# bar graph of GDP over time, coloured by continent
ggplot(gapm, aes(x = year, y = gdpPercap, fill = continent)) +
      geom_bar(stat = 'summary', fun.y = median) +
      scale_fill_brewer(type = 'qual')

# scatterplots of life expectancy by year, coloured by GDP, 
# facetted by continent
ggplot(gapm, aes(x = year, y = lifeExp, color = log10(gdpPercap))) +
      geom_jitter(size = 3) +
      facet_wrap(~ continent)

# line graph of change in life expectancy by country, faceted by
# continent
ggplot(subset(gapm, year == 1952 | year == 2007), aes(x = year, y = lifeExp)) +
  geom_line(aes(color = country)) +
  scale_colour_manual(values = country_colors) +
      facet_wrap(~ continent)

# line graphs of median life expectancy, GDP and population
# over time, by continent
p1 <- ggplot(gapm, aes(x = year, y = lifeExp, color = continent)) +
            geom_line(stat = 'summary', fun.y = median)

p2 <- ggplot(gapm, aes(x = year, y = gdpPercap, color = continent)) +
            geom_line(stat = 'summary', fun.y = median)

p3 <- ggplot(gapm, aes(x = year, y = pop, color = continent)) +
            geom_line(stat = 'summary', fun.y = median)

grid.arrange(p1, p2)

```

